|

### 2021-12-05 01-Chp-Background.Rmd
Working version

* Inclusion of preprints
* Individual patient data meta-analysis
* Triangulation across evidence sources

@@

### Individual patient data meta-analysis

@@

-   **Chapter \@ref(background-heading):** Background information on dementia and blood lipid levels. This chapter provides an introduction to the topics covered in this thesis to non-subject area experts, and discusses the motivation for the remainder of the thesis.
-   **Chapter \@ref(sys-rev-tools-heading):** This chapter introduces a new tool, `medrxivr`, which was used to developed to allow for systematic searches of the health-related preprint repositories.
-   **Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading):** These chapters describe, respectively, the methods and results of a comprehensive systematic review and meta-analysis. This review examined all available evidence on the effect of blood lipids and interventions that modified blood lipids on dementia outcomes.
-   **Chapter \@ref(cprd-analysis-heading):** This chapter examines the relationship between lipid-regulating agent use and dementia outcomes in the Clinical Practice Research Datalink, a large primary care electronic health record database.
-   **Chapter \@ref(ipd-heading):** This chapter describes an individual patient data analysis of several longitudinal cohort studies to describe the relationship between blood serum lipids and dementia outcomes, stratified by important covariates such as sex.
-   **Chapter \@ref(tri-heading)**: This chapter integrates the evidence identified and produced by the preceeding chapters as part of a quantitative triangulation framework.
-   **Chapter \@ref(discussion-heading)**: This chapter summarise the key clinical and methodological findings of the thesis and discusses their implications for practice. The overall strengths and weaknesses of this project are discussed in detail, and further avenues of research are suggested.

>>>>

* Inclusion of preprints
* Individual participant data meta-analysis
* Triangulation across evidence sources

@@

### Individual participant data meta-analysis

@@

-   **Chapter \@ref(background-heading):** Background information on dementia and blood lipid levels. This chapter provides an introduction to the topics covered in this thesis to non-subject area experts, and discusses the motivation for the remainder of the thesis.
-   **Chapter \@ref(sys-rev-tools-heading):** This chapter introduces a new tool, `medrxivr`, which was used to developed to allow for systematic searches of the health-related preprint repositories.
-   **Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading):** These chapters describe, respectively, the methods and results of a comprehensive systematic review and meta-analysis. This review examined all available evidence on the effect of blood lipids and interventions that modified blood lipids on dementia outcomes.
-   **Chapter \@ref(cprd-analysis-heading):** This chapter examines the relationship between lipid-regulating agent use and dementia outcomes in the Clinical Practice Research Datalink, a large primary care electronic health record database.
-   **Chapter \@ref(ipd-heading):** This chapter describes an individual participant data analysis of several longitudinal cohort studies to describe the relationship between blood serum lipids and dementia outcomes, stratified by important covariates such as sex.
-   **Chapter \@ref(tri-heading)**: This chapter integrates the evidence identified and produced by the preceeding chapters as part of a quantitative triangulation framework.
-   **Chapter \@ref(discussion-heading)**: This chapter summarise the key clinical and methodological findings of the thesis and discusses their implications for practice. The overall strengths and weaknesses of this project are discussed in detail, and further avenues of research are suggested.

### 2021-12-05 04-Chp-Systematic-Review-Results.Rmd
End of day:
Working version

One preprint (So _et al_[@so2017]) provided additional evidential value in a single meta-analysis (`r comma(1/18*100)`% of the 18 meta-analyses performed in this chapter). This meta-analysis of Mendelian randomisation studies examined the effects of lipid-lowering via HMGCR mutations on incidence of Alzheimer's disease (Figure \@ref(fig:mrStatinADFig)). To assess the evidence added to the meta-analysis through inclusion of the preprint, the data was re-analysed using a fixed-effect model. Examination of the weight assigned to each result in the analysis illustrates that a large proportion (`r mrStatinPreprintWeight`) of the weight is given to the preprinted result. However, the inclusion of the preprinted study did not have a substantial impact on the result of the meta-analysis (`r mrStatin$AD$estimate`) as compared to that reported by the single published study (`r estimate(0.59,0.25,1.39,"RR")`).

>>>>

One preprint (So _et al_[@so2017]) provided additional evidential value in a single meta-analysis (`r comma(1/18*100)`% of the 18 meta-analyses performed in this chapter). This meta-analysis of Mendelian randomisation studies examined the effects of lipid-lowering via HMGCR mutations on incidence of Alzheimer's disease (Figure \@ref(fig:mrStatinADFig)). To assess the evidence added to the meta-analysis through inclusion of the preprint, the data was re-analysed using a fixed-effect model. Examination of the weight assigned to each result in the analysis illustrates that a large proportion (`r mrStatinPreprintWeight`) of the weight is given to the preprinted result. However, the inclusion of the preprinted study did not have a substantial impact on the result of the meta-analysis (`r mrStatin$AD$estimate`) compared with that reported by the single published study (`r estimate(0.59,0.25,1.39,"RR")`), other than a slight increase in precision.

### 2021-12-05 06-Chp-IPD-Analysis.Rmd
Working version

@@

@@

In hindsight, attempting to undertake a large-scale IPD meta-analysis as part of a larger PhD project may have been overly ambitious. Data harmonization between cohorts in an IPD analysis is an often under-appreciated challenge,[@levis2021] and in line with this, data cleaning for this analysis took substantially longer than expected. While the cohort response rate was substantially lower that expected, given the time and resource of data cleaning and harmonisation for just three cohorts, a situation in which all 37 cohorts responded positively would have been logistically challenging within the scope of my PhD.

@@

\newpage

## References

>>>>

<!-- At the same time, I will write to authors to obtain anonymised data on individual participants. The data from each study will be combined using a technique known as individual participant data (IPD) meta-analysis. This will allow me to examine if the relationship between lipids in dementia depends on any patient characteristics (e.g. sex or age). -->

@@

<!-- TODO May need to talk about other confounders here -->

@@

In hindsight, attempting to undertake a large-scale IPD meta-analysis as part of a larger PhD project may have been overly ambitious. Data harmonization between cohorts in an IPD analysis is an often under-appreciated challenge,[@levis2021] and in line with this, data cleaning for this analysis took substantially longer than expected. While the cohort response rate was substantially lower that expected, given the time and resource of data cleaning and harmonisation for just three cohorts, a situation in which all 37 cohorts responded positively would have been logistically challenging within the scope of a PhD.

@@

### 2021-12-05 07-Chp-Triangulation.Rmd
Working version

```{block type='savequote', include=knitr::is_latex_output(), quote_author='(ref:tri-quote)', echo = TRUE}
Next on my list of features to be specially considered I would place the consistency of the observed association. Has it been repeatedly observed by different persons, in different places, circumstances and times?
```
(ref:tri-quote) --- Sir Austin Bradford Hill, 1965 [@hill1965]

@@

## Qualitative (narrative) triangulation

@@

In the following section, I propose and apply a novel quantitative triangulation framework to address these limitations. The absence of any clear signals across the evidence base does not make for particularly interesting case studies but I will nonetheless use the identified evidence to illustrate the novel methodological approach.

@@

In addition to the qualitative discussion of the evidence presented in Section (\@ref(qual-tri)), I attempted to integrate the numerical results of the multiple approaches under a proposed quantitative triangulation framework. This approach incorporates recent advancements in the way that bias in results is assessed (namely the move to domain-based assessment tools)[@sterne2019] and existing methods for bias-/indirectness-adjusted meta-analysis[@turner2009] to illustrate how causal questions could be addressed under this framework.

This proposed framework involves several steps, described in detail in the following sections. In summary, these steps are:

1. Define the causal question of interest
1. Identify of relevant evidence sources and standardise effect directions
1. Specify idealised version of each study
1. Assess and visualise the extent and direction of bias in each result
1. Define modifying terms for bias and indirectness in each result
1. Calculate bias-/indirectness-adjusted results and perform meta-analysis

@@

__Case study #1: Effect of mid-life LDL-c on Alzheimer's disease__

@@

__Case study #2: Effect of mid-life triglycerides on vascular dementia__

@@

@@

These questions were chosen for several reasons. Firstly, work presented in previous chapters had identified some evidence for an association in these exposure/outcome pairs. Secondly, given the long prodomal period between the onset of physiological changes in the brain and presentation of dementia symptoms, it seems likely that conditions during the mid-life period are particularly important. Finally, examining the causal impact of a lipid faction on a specific outcome (e.g. Alzheimer's disease), rather than a composite outcome (e.g. all-cause dementia), focuses on an single causal pathway rather than including several conditions in the outcome definition that likely have very different mechanisms of disease. In addition, previous work presented in this thesis illustrated that the component conditions (in this case Alzheimer's disease and vascular dementia) are likely to be subject to very different confounding structures (see Section \@ref(cprd-confounding-by-ind)), and use of an composite all-cause outcome in the triangulation framework may mask this.

@@

@@

####	Assign modifying values to risk/direction of bias
<!-- TODO Note that they are on the log scale - Lines 45-50 of addbias -->

@@

<!-- QUESTION sign is negaitve if to the left, because you are taking it away in the final formula and it therefore needs a negative sign to be positive overall! -->

@@

@@

#### Accuracy of bias-/indirectness-assessment

@@

>>>>

```{block type='savequote', include=knitr::is_latex_output(), quote_author='(ref:tri-quote)', echo = TRUE}
Next on my list of features to be specially considered I would place the consistency of the observed association. Has it been repeatedly observed by different persons, in different places, circumstances and times?
```
(ref:tri-quote) --- Austin Bradford Hill, 1965 [@hill1965]

@@

## Qualitative (narrative) triangulation {#qual-tri}

@@

In the following section, I propose and apply a novel quantitative triangulation framework to address these limitations. This approach incorporates recent advancements in the way that bias in results is assessed (namely the move to domain-based assessment tools)[@sterne2019] and existing methods for bias-/indirectness-adjusted meta-analysis[@turner2009] to integrate the numerical results of the multiple approaches. The absence of any clear signals across the evidence base, as detailed in the previous section, does not make for particularly interesting case studies but I will nonetheless use the identified evidence to illustrate the novel methodological approach.

@@

The proposed framework involves several steps, described in detail in the following sections. In summary, these steps are:

1. Define the causal question of interest
1. Identify of relevant evidence sources and standardise effect directions
1. Specify idealised version of each study
1. Assess the extent and direction of bias/indirectness in each result
1. Define modifying terms for bias and indirectness in each result
1. Calculate bias-/indirectness-adjusted results and perform meta-analysis

@@

_Case study #1: Effect of mid-life LDL-c on Alzheimer's disease_

@@

_Case study #2: Effect of mid-life triglycerides on vascular dementia_

@@

<!-- QUESTION Should population here include a geographical aspect, as in the Discussion I say that the generalisability to other populations is an issue. -->

@@

These causal questions were chosen for several reasons. Firstly, work presented in previous chapters had identified some evidence for an association in these exposure/outcome pairs. Secondly, given the long prodomal period between the onset of physiological changes in the brain and presentation of dementia symptoms, it seems likely that conditions during the mid-life period are particularly important. Finally, examining the causal impact of a lipid faction on a specific outcome (e.g. Alzheimer's disease), rather than a composite outcome (e.g. all-cause dementia), focuses on an single causal pathway rather than including several conditions in the outcome definition that likely have very different mechanisms of disease. In addition, previous work presented in this thesis illustrated that the component conditions (in this case Alzheimer's disease and vascular dementia) are likely to be subject to very different confounding structures (see Section \@ref(cprd-confounding-by-ind)), and use of an composite all-cause outcome in the triangulation framework may mask this.

@@

<!-- TODO Note that the results are converted to the log scale. -->

@@

####	Assign modifying values to risk/direction of bias
<!-- TODO Note that they are on the log scale, and that this requires the results to be on the log scale too! - Lines 45-50 of addbias -->

@@

<!-- QUESTION sign is negative if to the left, because you are taking it away in the final formula and it therefore needs a negative sign to be positive overall! -->

<!-- TODO Need to explain here that for proportional bias, the results are exponentiated  -->

@@

<!-- TODO Talk about validation at some point -->

@@

#### Accuracy of bias-/indirectness-assessment {#tri-accuracy}

@@

<!-- TODO An extra step of the process should be the consideration of whether it is reasonable to synthesis across different effect estimates. Fine for rare, not for common. Fancy models that can be used to synthesis effect estimates are still compatabile with this approach as each invidual estimate is first adjusted and the adjusted estimate can then be used with whatever meta-anlaysis technique you like. -->
